Temporal graph realization from fastest paths

### 2024-05-22 Arxiv-30-4-24/ARXIV+nina-corrections.tex
Added changes from Nina's PhD thesis comments

We remark that a fastest path between two vertices in a temporal graph can be computed in polynomial time~\cite{xuan_computing_2003,Wu2016Efficient}.
%Wu et~al.~\cite{Wu2016Efficient} provide an algorithm that computes the duration of the fastest temporal path from a source vertex $v$, in a temporal graph $(G,\lambda)$, to all other vertices,
%that runs in polynomial-time.
Hence, given a $\Delta$-periodic temporal graph $(G,\lambda,\Delta)$, we can compute in polynomial-time  the matrix $D$
which consists of durations of fastest temporal paths among all pairs of vertices in $(G,\lambda,\Delta)$.

@@

\begin{proof}
	We present a polynomial-time reduction from the NP-hard problem NAE 3-SAT~\cite{Schaefer1978complexity}. Here we are given a formula $\phi$ that is a conjunction of so-called NAE (not-all-equal) clauses, where each clause contains exactly 3 literals (with three distinct variables).
    A NAE clause evaluates to \textsc{true} if and only if not all of its literals are equal, that is, at least one literal evaluates to \textsc{true} and at least one literal evaluates to \textsc{false}.
    We are asked whether $\phi$ admits a satisfying assignment.

@@

%
\begin{figure}[t]
	\noindent
	\makebox[\textwidth]{
		\centering
	\includegraphics{fig-NP-reduction}
}
\caption{Illustration of the temporal graph $(G,\lambda)$ from the NP-hardness reduction, 
	where the NAE 3-SAT formula $\phi$ is of the form $\phi = \text{NAE}(x_1, \overline{x}_2, x_3) \wedge \text{NAE}(x_1, x_2, x_4)$.
To improve the readability, we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with gray dotted lines.
Presented is the labeling of $G$ corresponding to the assignment $x_1=x_2=\textsc{true}$ and $x_3,x_4=\textsc{false}$,
where all unlabeled edges get the label $2$.
}\label{fig:NP-example}
\end{figure}
%
We set $\Delta$ to some constant larger than two, that is, $\Delta\ge 3$. Next, we specify the durations in the matrix $D$ between all vertex pairs.
For the sake of simplicity we write $D_{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following durations:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following durations:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following duratios:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following duratios:
    $d(x_i,c)=2 \Delta$, $d(c,x_i)=\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following duratios:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following duratios:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of the instance $(D,\Delta)$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $(D,\Delta)$ is a \textsc{Yes}-instance of \deltaExact\ if and only if NAE 3-SAT formula $\phi$ is satisfiable.

@@

We construct a labeling $\lambda$ as follows.
\begin{itemize}
    \item All edges incident with a clause vertex $c$ obtain label one.
    %\item For each variable $x$, we set $\lambda(\{x^T,x^F\})=1$.
    \item If variable $x_i$ is set to \textsc{true}, we set $\lambda(\{x_i^F, v\})=3$.
    \item If variable $x_i$ is set to \textsc{false}, we set $\lambda(\{x_i^T, v\})=3$.
    \item We set the labels of all other edges to two.
\end{itemize}
For an example of the constructed temporal graph see~\cref{fig:NP-example}.
We now verify that all duratios are realized.
\begin{itemize}
    \item For each variable $x_i$ in $\phi$ we have to check that $d(x_i,v)=2$
    and $d(v,x_i)=\Delta$.

@@

    Let $x_k$ be a variable that appears non-negated in $c_i$
    and $x_\ell$ the variable that appears non-negated in $c_j$
    (all other cases are symmetric).
    There is a path of length three from $c_i$ to $c_j$
    that passes first through vertex $x_k^T$ and then through vertex $x_\ell^T$.
    Therefore the temporal path from $c_i$ to $c_j$
    uses the edges $\{x_k^T, c_i\}$, $\{x_\ell^T, c_j\}$, and $\{x_k^T, x_\ell ^T\}$, with labels $1$,~$2$, and $1$ (at times $1$,~$2$, and $\delta + 1$), respectively,
    which proves the desired equality.
    Note also that this is the shortest path between $c_i$ and $c_j$, and that the first and the last edge must have the label $1$,
    therefore it follows that this is the fastest temporal path.
\end{itemize}
Lastly, observe that the above constructed labeling $\lambda$ uses values $\{1,2,3\}\subseteq[\Delta]$, therefore $\Delta\ge 3$.
\end{proof}

@@

\begin{proof}
    Let $D$ be an input matrix for problem \deltaExact\ of dimension $n \times n$.
    Let us fix the vertices of the corresponding graph $G$ of $D$ as $v_1, v_2, \dots, v_n$, where vertex $v_i$ corresponds to the row and column $i$ of matrix $D$.
    This can be done in polynomial time as we need to loop through the matrix $D$ once and connect vertices $v_i, v_j$ for which $D_{i,j} = 1$. At the same time we also check if $D_{i,i} = 0$, for all $i \in [n]$.
    When $G$ is constructed we run DFS algorithm on it and check that it has no cycles.
    If at any step we encounter a problem, our algorithm stops and returns a negative answer.

    Having computed $G$, our algorithm proceeds as follows. We pick an arbitrary edge $f$ and give it label one, that is, $\lambda(f)=1$. Now we push all edges incident with $f$ into a (initially empty) queue. Now we repeat the following as long as the queue is not empty:
\begin{itemize}
    \item Pop edge $e=\{u,v\}$ from the queue. Since $e$ was pushed into the queue, there is an edge $e'$ incident with $e$ that already obtained a label. Let w.l.o.g.\ $e'=\{v,w\}$. Then we set $\lambda(e)=(\lambda(e')-D_{u,w}+1)\bmod \Delta$.
    \item Push all edges incident with $e$ that have not received a label yet into the queue.
\end{itemize}
When the queue is empty, all edges have received a label. Iterate over all vertex pairs $u,v$ and check whether the fastest path from $u$ to $v$ in $(G,\lambda)$ has duration $D_{u,v}$. If this check succeeds for all vertex pairs, output the labeling $\lambda$, otherwise abort.

@@

$(\Leftarrow)$: Assume we face a yes-instance, then there exists a labeling $\lambda^\star$ that realizes all durations specified in $D$. Let $e^\star$ denote the edge initially picked by the algorithm. For all edges $e$ let $\lambda(e)=(\lambda^\star(e)-\lambda^\star(e^\star)+1)\bmod \Delta$. Clearly, the labeling $\lambda$ also realizes all durations specified in $D$ since $\lambda$ is obtained by adding the constant $(1-\lambda^\star(e^\star))$ modulo $\Delta$ to all labels of $\lambda^\star$ which does not change the duration of any temporal path, that is all durations in $(G,\lambda^\star)$ are the same as their counterparts in $(G,\lambda)$. We claim that our algorithm computes and outputs $\lambda$.

@@

Initially, our algorithm labels $e^\star$ with one, which equals $\lambda(e^\star)$. Now let $e$ be an edge popped off the queue by the algorithm in some iteration, that is on the distance $i$ from $e^\star$. Let $e'$ be the edge incident with $e$ that is on the distance $i-1$ from $e^\star$. Since $G$ is a tree $e'$ has already been considered by the algorithm and thus already has a label. By induction we have that the algorithm labeled $e'$ with $\lambda(e')$. 
Assume that $e=\{u,v\}$ and $e'=\{v,w\}$. Since $G$ is a tree there is only one path from $u$ to $w$ in $G$ and it uses edges $e$ and $e'$. It follows that $\lambda(e')-\lambda(e)+1=D_{u,w}$ if $\lambda(e')>\lambda(e)$, and $\lambda(e')-\lambda(e)+\Delta+1=D_{u,w}$ otherwise. Our algorithm labels $e$ with $(\lambda(e')-D_{u,w}+1)\bmod \Delta$. It is straightforward to verify that the label of $e$ computed by the algorithm equals $\lambda(e)$. It follows that the algorithm computes~$\lambda$.
\end{proof}

@@

%Let us remember the following basic property of the problem \deltaExact.
Recall that a labeling $\lambda$ of $G$ satisfies $D$ if the duration of a fastest temporal path from each vertex $v_i$ to each other vertex $v_j$ equals $D_{v_i, v_j}$.
%, and no other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases.
We consider the fastest temporal paths 
where the starting vertex is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the destination vertex is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases, we guess the underlying path $P$ that at least one fastest temporal path from the vertex $v_i$ to $v_j$ follows, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $f(k) \cdot |D| ^ {O(1)}$ constraints. 
Note that we have to do this while keeping the total number of variables upper-bounded by some function in $k$.
%We have to guarantee only that the number of guesses of fastest temporal paths is exponential in $k$ and not in $n$.

@@

    We want to guess a split vertex of $w$ in $S_{u,v}$, and the fastest temporal path that reaches it.

@@

			\includegraphics{fig-guess2}}
		\caption{Example of an Inner segment guess II (\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT:guess-uToSegmentz2}}}}), where we guessed the fastest temporal paths of the form $w \leadsto u \rightarrow v_2$ (in blue) and $w \leadsto v \rightarrow v_{p-1}$ (in red). 
			\label{fig:FPT-guessG5}}
	\end{subfigure}

@@

    \begin{definition}[Travel delays] \label{def:travel-delay}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of \deltaExact.
    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
    More specifically:
    \begin{equation}\label{eq:def-VertexWaitingTime}
       \tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
    \end{equation}
    Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
    \end{definition}
    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

@@

Suppose now that 
we know the underlying path $P_{u,v} = (u=v_1, v_2, \dots, v_p = v)$ of the fastest temporal path between vertices of interest $u$ and $v$ in $G'$.
Let $v_i\in U$ with $u\neq v_i \neq v$ be a vertex of interest on the path $P_{u,v}$.
Suppose that $v_i$ is reached the fastest from $u$ by a path $P = (u = u_1, u_2, \dots, u_{j-1}, v_i)$.
We split the path with $P_{u,v}$ into a path $ Q = (u=v_1, v_2, \dots, v_i)$ and $R = (v_i, v_{i+1}, \dots, v_p=v)$
(for details see~\cref{fig:FPT-uv-Labelalledges}).

@@

\begin{figure}[t]
	\centering
	\includegraphics{fig-lemma}
	\caption{An example of the situation in \cref{lemma:FPT-uv-LabelAlmostalledges},
		where we assume that the fastest temporal path from $u$ to $v$ is $P_{u,v} = (u=v_1, v_2, \dots v_p)$,
		and the fastest temporal path from $u$ to $v_i$ in $P_{u,v}$ is $P = (u, u_2, u_3, \dots, v_i)$.
		We denote with $Q = (u=v_1, v_2, \dots, v_i)$ and with $R = (v_i, v_{i+1}, \dots, v_p=v)$.
		\label{fig:FPT-uv-Labelalledges}}
\end{figure}

@@

    \emph{Case 1: }$f = v_2v_3$.
    Let us observe the case of the fastest temporal paths between $v_{1}$ and $v_{3}$.
    Denote with $Q = (v_{1}, v_2, v_3)$ and with $P' = (v_3,v_4,v_5,v_6)$.
    From our proposition, it follows that
    \begin{itemize}
        \item the fastest temporal path $P ^ +$ from $v_1$ to $v_3$ 
    is of the following form 
    $P^+ = v_{1}  \leadsto v_6 \rightarrow v_5 \rightarrow v_4 \rightarrow v_3$,
    and
        \item the fastest temporal path $P ^ -$ from $v_{3}$ to $v_{1}$ 
    is of the following form 
    $P^- = v_{3} \rightarrow v_{4} \rightarrow v_5 \rightarrow v_6 \leadsto v_1$.
    \end{itemize}
    It follows that 
    $
    d(v_{1}, v_{3}, P^+) \leq d(v_{1}, v_{3}, Q),
    $
    and
    $
    d(v_{1}, v_{3}, P^-) \leq d(v_{1}, v_{3}, Q)
    $. 
    Note that $d(v_{1}, v_{3}, P^+) \geq 1 + d(v_6,v_3,P')$,
    and by the definition $d(v_6,v_3,P') = 1 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta$,
    where $(t_i - t_j)_\Delta$ denotes the difference of two consecutive labels $t_i, t_j$ modulo $\Delta$.
    Similarly holds for $d(v_{1}, v_{3}, P^+)$.
    Summing now both of the above equations we get
    \begin{equation}
    \begin{split} \label{eq:FPT-prf-fUnlabeled}
        d(v_{1}, v_{3}, P^+) + d(v_{3}, v_{1}, P^-) &\leq 
        d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        1 + d(v_6,v_3,P') + 1 + d(v_3,v_6,P') &\leq d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        3 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta +
        1 + (t_4 - t_3)_\Delta + (t_5 - t_4)_\Delta 
        &\leq 
        1 + (t_2 - t_1)_\Delta + 
        1 + (t_1 - t_2)_\Delta\\
        %
        (t_4 - t_5)_\Delta + (t_5 - t_4)_\Delta +
        (t_4 - t_3)_\Delta + (t_3 - t_4)_\Delta  + 2 
        &\leq 
        (t_2 - t_1)_\Delta + (t_1 - t_2)_\Delta.
    \end{split}
    \end{equation}
    Note that if $t_i \neq t_j$ we get that 
    the sum
    $(t_i - t_j)_\Delta + (t_j - t_i)_\Delta$ equals exactly $\Delta$,
    and if $t_i = t_j$ the sum equals $2\Delta$.
    This follows from the definition of travel delays at vertices (see \cref{obs:travel-delays-both-directions}).
    Therefore we get from \cref{eq:FPT-prf-fUnlabeled}, 
    that the right part is at most $2 \Delta$, while the left part is at least $2 \Delta + 1$,
    for any relation of labels $t_1,t_2, \dots, t_5$,
    which is a contradiction.

>>>>

We remark that a fastest path between two vertices in a temporal graph can be computed in polynomial time~\cite{xuan_computing_2003,Wu2016Efficient}.
%Wu et~al.~\cite{Wu2016Efficient} provide an algorithm that computes the duration of the fastest temporal path from a source vertex $v$, in a temporal graph $(G,\lambda)$, to all other vertices,
%that runs in polynomial-time.
Hence, given a $\Delta$-periodic temporal graph $(G,\lambda,\Delta)$, we can compute in polynomial time  the matrix $D$
which consists of durations of fastest temporal paths among all pairs of vertices in $(G,\lambda,\Delta)$.

@@

\begin{proof}
	We present a polynomial-time reduction from the NP-hard problem NAE 3-SAT~\cite{Schaefer1978complexity}. We are given a formula $\phi$ that is a conjunction of so-called NAE (not-all-equal) clauses, where each clause contains exactly 3 literals (with three distinct variables).
    A NAE clause evaluates to \textsc{true} if and only if not all of its literals are equal, that is, at least one literal evaluates to \textsc{true} and at least one literal evaluates to \textsc{false}.
    We are asked whether $\phi$ admits a satisfying assignment.

@@

%
\begin{figure}[t]
	\noindent
	\makebox[\textwidth]{
		\centering
	\includegraphics{fig-NP-reduction}
}
\caption{Illustration of the temporal graph $(G,\lambda)$ from the NP-hardness reduction, 
	where the NAE 3-SAT formula $\phi$ is of the form $\phi = \text{NAE}(x_1, \overline{x}_2, x_3) \wedge \text{NAE}(x_1, x_2, x_4)$.
To improve the readability, we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with gray dotted lines.
Presented is the labeling of $G$ corresponding to the assignment $x_1=x_2=\textsc{true}$ and $x_3,x_4=\textsc{false}$,
where all unlabeled edges get the label $2$.
}\label{fig:NP-example}
\end{figure}
%
We set $\Delta$ to some constant larger than two, that is, $\Delta\ge 3$. Next, we specify the durations in the matrix $D$ between all vertex pairs.
For the sake of simplicity we write $D_{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following durations:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following durations:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following durations:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following durations:
    $d(x_i,c)=2 \Delta$, $d(c,x_i)=\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following durations:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following durations:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of the instance $(D,\Delta)$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $(D,\Delta)$ is a \textsc{Yes}-instance of \deltaExact\ if and only if NAE 3-SAT formula $\phi$ is satisfiable.

@@

We construct a labeling $\lambda$ as follows.
\begin{itemize}
    \item All edges incident with a clause vertex $c$ obtain label one.
    %\item For each variable $x$, we set $\lambda(\{x^T,x^F\})=1$.
    \item If variable $x_i$ is set to \textsc{true}, we set $\lambda(\{x_i^F, v\})=3$.
    \item If variable $x_i$ is set to \textsc{false}, we set $\lambda(\{x_i^T, v\})=3$.
    \item We set the labels of all other edges to two.
\end{itemize}
For an example of the constructed temporal graph see~\cref{fig:NP-example}.
We now verify that all durations are realized.
\begin{itemize}
    \item For each variable $x_i$ in $\phi$ we have to check that $d(x_i,v)=2$
    and $d(v,x_i)=\Delta$.

@@

    Let $x_k$ be a variable that appears non-negated in $c_i$
    and $x_\ell$ the variable that appears non-negated in $c_j$
    (all other cases are symmetric).
    There is a path of length three from $c_i$ to $c_j$
    that passes first through vertex $x_k^T$ and then through vertex $x_\ell^T$.
    Therefore the temporal path from $c_i$ to $c_j$
    uses the edges $\{x_k^T, c_i\}$, $\{x_\ell^T, c_j\}$, and $\{x_k^T, x_\ell ^T\}$, with labels $1$,~$2$, and $1$ (at times $1$,~$2$, and $\Delta + 1$), respectively,
    which proves the desired equality.
    Note also that this is the shortest path between $c_i$ and $c_j$, and that the first and the last edge must have the label $1$,
    therefore it follows that this is the fastest temporal path.
\end{itemize}
Lastly, observe that the above constructed labeling $\lambda$ uses values $\{1,2,3\}\subseteq[\Delta]$, therefore $\Delta\ge 3$.
\end{proof}

@@

\begin{proof}
    Let $D$ be an input matrix of \deltaExact\ of dimension $n \times n$.
    Let us fix the vertices of the corresponding graph $G$ of $D$ as $v_1, v_2, \dots, v_n$, where vertex $v_i$ corresponds to the row and column $i$ of matrix $D$.
    This can be done in polynomial time as we need to loop through the matrix $D$ once and connect vertices $v_i, v_j$ for which $D_{i,j} = 1$. At the same time we also check if $D_{i,i} = 0$, for all $i \in [n]$.
    When $G$ is constructed we run DFS algorithm on it and check that it has no cycles.
    If at any step we encounter a problem, our algorithm stops and returns a negative answer.

    Having computed $G$, our algorithm proceeds as follows. We pick an arbitrary edge $f$ and give it label one, that is, $\lambda(f)=1$. Now we push all edges incident with $f$ into a (initially empty) queue. Now we repeat the following as long as the queue is not empty:
\begin{itemize}
    \item Pop edge $e=\{u,v\}$ from the queue. Since $e$ was pushed into the queue, there is an edge $e'$ incident with $e$ that already obtained a label. Let w.l.o.g.\ $e'=\{v,w\}$. Then we set $\lambda(e)=(\lambda(e')-D_{u,w}+1)\bmod \Delta$.
    \item Push all edges incident with $e$ that have not received a label yet into the queue.
\end{itemize}
When the queue is empty, all edges have received a label. Iterate over all vertex pairs $u,v$ and check whether the fastest path from $u$ to $v$ in $(G,\lambda)$ has duration $D_{u,v}$. If this check succeeds for all vertex pairs, output yes, otherwise abort.

@@

$(\Leftarrow)$: Assume we face a yes-instance, then there exists a labeling $\lambda^\star$ that realizes all durations specified in $D$. Let $e^\star$ denote the edge initially picked by our algorithm. For all edges $e$ let $\lambda(e)=(\lambda^\star(e)-\lambda^\star(e^\star)+1)\bmod \Delta$. Clearly, the labeling $\lambda$ also realizes all durations specified in $D$ since $\lambda$ is obtained by adding the constant $(1-\lambda^\star(e^\star))$ modulo $\Delta$ to all labels of $\lambda^\star$ which does not change the duration of any temporal path, that is all durations in $(G,\lambda^\star)$ are the same as their counterparts in $(G,\lambda)$. We claim that our algorithm computes and outputs $\lambda$.

@@

Initially, our algorithm labels $e^\star$ with $1$, which equals $\lambda(e^\star)$. Now let $e$ be an edge popped off the queue by the algorithm in some iteration, that is on the distance $i$ from $e^\star$. Let $e'$ be the edge incident with $e$ that is on the distance $i-1$ from $e^\star$. Since $G$ is a tree $e'$ has already been considered by the algorithm and thus already has a label. By induction we have that the algorithm labeled $e'$ with $\lambda(e')$. 
Assume that $e=\{u,v\}$ and $e'=\{v,w\}$. Since $G$ is a tree there is only one path from $u$ to $w$ in $G$ and it uses edges $e$ and $e'$. It follows that $\lambda(e')-\lambda(e)+1=D_{u,w}$ if $\lambda(e')>\lambda(e)$, and $\lambda(e')-\lambda(e)+\Delta+1=D_{u,w}$ otherwise. Our algorithm labels $e$ with $(\lambda(e')-D_{u,w}+1)\bmod \Delta$. It is straightforward to verify that the label of $e$ computed by the algorithm equals $\lambda(e)$. It follows that the algorithm computes~$\lambda$.
\end{proof}

@@

%Let us remember the following basic property of the problem \deltaExact.
Recall that a labeling $\lambda$ of $G$ satisfies $D$ if the duration of a fastest temporal path from each vertex $v_i$ to each other vertex $v_j$ equals $D_{v_i, v_j}$.
%, and no other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases.
We consider the fastest temporal paths 
where the starting vertex is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the destination vertex is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases, we guess the underlying path $P$ that at least one fastest temporal path from the vertex $v_i$ to $v_j$ follows, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $f(k) \cdot |D| ^ {O(1)}$ constraints, where $|D|= n^2$. 
Note that we have to do this while keeping the total number of variables upper-bounded by some function in $k$.
%We have to guarantee only that the number of guesses of fastest temporal paths is exponential in $k$ and not in $n$.

@@

    We want to determine the split vertex of $w$ in $S_{u,v}$, and guess the fastest temporal path that reaches it.

@@

			\includegraphics{fig-guess2}}
		\caption{Example of an Inner segment guess II (\textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT:guess-uToSegmentz2}}}}), where we guessed the fastest temporal paths of the form $w \leadsto u \rightarrow v_2$ (in blue) and $w \leadsto v \rightarrow  v_{p-1} \rightarrow v_2$ (in red). 
			\label{fig:FPT-guessG5}}
	\end{subfigure}

@@

    \begin{definition}[Travel delays] \label{def:travel-delay}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of \deltaExact.
    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$, if the labels are different, or we set it to $\Delta$.
    More specifically:
    \begin{itemize}
        \item if $\lambda(e_1) \neq \lambda(e_2)$ we have
    \begin{equation}\label{eq:def-VertexWaitingTime}
       \tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta,
    \end{equation}
    \item if $\lambda(e_1) = \lambda(e_2)$ we have
    \begin{equation*}
    \tau_v^{uz} = \Delta.
    \end{equation*}
    \end{itemize}
    %Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
    \end{definition}
    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

@@

Suppose now that 
we know the underlying path $P_{u,v} = (u=v_1, v_2, \dots, v_p = v)$ of the fastest temporal path between vertices of interest $u$ and $v$ in $G'$.
Let $v_i\in U$ with $u\neq v_i \neq v$ be a vertex of interest on the path $P_{u,v}$.
Suppose that $v_i$ is reached the fastest from $u$ by a path $P = (u = u_1, u_2, \dots, u_{j-1}, v_i)$.
We split the path $P_{u,v}$ into a path $ Q = (u=v_1, v_2, \dots, v_i)$ and $R = (v_i, v_{i+1}, \dots, v_p=v)$
(for details see~\cref{fig:FPT-uv-Labelalledges}).

@@

\begin{figure}[t]
	\centering
	\includegraphics{fig-lemma}
	\caption{An example of the situation in \cref{lemma:FPT-uv-LabelAlmostalledges},
		where we assume that the fastest temporal path from $u$ to $v$ is $P_{u,v} = (u=v_1, v_2, \dots v_p)$,
		and the fastest temporal path from $u$ to $v_i$ is $P = (u, u_2, u_3, \dots, v_i)$.
		We denote $Q = (u=v_1, v_2, \dots, v_i)$ and $R = (v_i, v_{i+1}, \dots, v_p=v)$.
		\label{fig:FPT-uv-Labelalledges}}
\end{figure}

@@

    \emph{Case 1: }$f = v_2v_3$.
    Let us observe the case of the fastest temporal paths between $v_{1}$ and $v_{3}$.
    Denote $Q = (v_{1}, v_2, v_3)$ and $P' = (v_3,v_4,v_5,v_6)$.
    From our proposition, it follows that
    \begin{itemize}
        \item the fastest temporal path $P ^ +$ from $v_1$ to $v_3$ 
    is of the following form 
    $P^+ = v_{1}  \leadsto v_6 \rightarrow v_5 \rightarrow v_4 \rightarrow v_3$,
    and
        \item the fastest temporal path $P ^ -$ from $v_{3}$ to $v_{1}$ 
    is of the following form 
    $P^- = v_{3} \rightarrow v_{4} \rightarrow v_5 \rightarrow v_6 \leadsto v_1$.
    \end{itemize}
    It follows that 
    $
    d(v_{1}, v_{3}, P^+) \leq d(v_{1}, v_{3}, Q),
    $
    and
    $
    d(v_{3}, v_{1}, P^-) \leq d(v_{3}, v_{1}, Q)
    $. 
    Note that $d(v_{1}, v_{3}, P^+) \geq 1 + d(v_6,v_3,P')$,
    and by the definition $d(v_6,v_3,P') = 1 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta$,
    where $(t_i - t_j)_\Delta$ denotes the difference of two consecutive labels $t_i, t_j$ modulo $\Delta$.
    Similarly holds for $d(v_{1}, v_{3}, P^+)$.
    Summing now both of the above equations we get
    \begin{equation}
    \begin{split} \label{eq:FPT-prf-fUnlabeled}
        d(v_{1}, v_{3}, P^+) + d(v_{3}, v_{1}, P^-) &\leq 
        d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        1 + d(v_6,v_3,P') + 1 + d(v_3,v_6,P') &\leq d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        3 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta +
        1 + (t_4 - t_3)_\Delta + (t_5 - t_4)_\Delta 
        &\leq 
        1 + (t_2 - t_1)_\Delta + 
        1 + (t_1 - t_2)_\Delta\\
        %
        (t_4 - t_5)_\Delta + (t_5 - t_4)_\Delta +
        (t_4 - t_3)_\Delta + (t_3 - t_4)_\Delta  + 2 
        &\leq 
        (t_2 - t_1)_\Delta + (t_1 - t_2)_\Delta.
    \end{split}
    \end{equation}
    Note that if $t_i \neq t_j$ we get that 
    the sum
    $(t_i - t_j)_\Delta + (t_j - t_i)_\Delta$ equals exactly $\Delta$,
    and if $t_i = t_j$ the sum equals $2\Delta$.
    This follows from the definition of travel delays at vertices (see \cref{obs:travel-delays-both-directions}).
    Therefore we get from \cref{eq:FPT-prf-fUnlabeled}, 
    that the right part is at most $2 \Delta$, while the left part is at least $2 \Delta + 1$,
    for any relation of labels $t_1,t_2, \dots, t_5$,
    which is a contradiction.
